Extended Data

### 2021-10-15 doc/manuscript/Dow_manuscript_Nature.Rmd
worked on discussion

Together, our results demonstrate that warmer spring temperatures in the temperate deciduous forests of eastern North America advance the phenology of tree stem growth but have little effect on annual woody productivity (Figs. `r schematic_summary`- `r tree_rings`).

The observed advance in the timing of stem growth milestones under warmer springs parallels phenological advances observed for leaf development (Fig. `r DOY_timing`)[@keenan_net_2014] and NEE [@keenan_net_2014].
However, inconsistent with leaf phenology and NEE, we observed that the length of the period of most active stem growth does not increase under warmer springs. 
This suggests that the growth benefit of warmer springs is compensated by negative effects later in the growing season, with two primary possible mechanisms.
First, stem growth, like leaf phenology[@zani_increased_2020], may be sink-limited, such that an advance in the start of growth ($DOY_25$) would lead to an earlier depletion of resources needed for growth resulting in an earlier end to the season of rapid growth ($DOY_{75}$).
Warm springs may also be associated with greater seasonal water stress, effectively canceling out any positive effects of an extended growing period[@buermann_widespread_2018].

The concept that stem growth is not enhanced under a longer growing season was supported by our tree-ring analysis, which showed that the effect of spring temperatures on annual growth is negligible for most species and locations, with at most modest influence for some species and locations (Fig. `r tree_rings`, `r EDF_tree_rings_TMIN`, `r EDT_tree_ring_summary`)[@alexander_potential_2019; @helcoski_growing_2019].
Rather, the primary effect of temperature on annual growth is a negative effect of $T_{max}$ during the peak growing season, which is most likely a signal of drought stress[@dorangeville_drought_2018].
*It is possible that an earlier start to the growing season may provide some buffer against drought later in the growing season[@wolf_warm_2016], potentially contributing to observed trends of declining drought sensitivity in the region[@maxwell_declining_2016;@helcoski_growing_2019; @zhang_drought_2021]. (Is this relevent?)*<!-- (Neil suggested that the zhang paper would be good here but i'm not sure I see how...)-->
<!-- *(ADD MORE ON DROUGHT HERE)*-->

>>>>

Together, our results demonstrate that warmer spring temperatures in the temperate deciduous forests of eastern North America advance the phenology of tree stem growth but have little effect on annual woody productivity (Figs. `r schematic_summary`- `r tree_rings`). 
The observed phenological advance in the start of stem growth under warmer springs parallels phenological advances observed for leaf development (Fig. `r DOY_timing`)[@keenan_net_2014] and NEE [@keenan_net_2014].
However, inconsistent with the concept that an earlier start to growth would increase annual woody productivity, we demonstrate that warmer springs advance the end to the period of most active stem growth and have negligible effect on total annual growth for most species and locations (Fig. `r tree_rings`).

This suggests that any growth benefit of warmer springs is compensated by negative effects later in the growing season, with two primary possible mechanisms.
First, the cessation of rapid stem expansion, which occurs mid-summer near the time of peak leaf greenness (`r EDF_growth_curves_all`), is likely driven by cues other than carbon limitation, such as daylength or sink limitation[@zani_increased_2020].
<!--potentially expand on sink limitation here-->

*First, stem growth, like leaf phenology[@zani_increased_2020], may be sink-limited, such that an advance in the start of growth ($DOY_25$) would lead to an earlier depletion of resources needed for growth resulting in an earlier end to the season of rapid growth ($DOY_{75}$).*
*Warm springs may also be associated with greater seasonal water stress, effectively canceling out any positive effects of an extended growing period[@buermann_widespread_2018].*

*Rather, the primary effect of temperature on annual growth is a negative effect of $T_{max}$ during the peak growing season, which is most likely a signal of drought stress[@dorangeville_drought_2018].*
